---
title: "Homework 3 Additional Code"
date: 2024-06-02
author: "Jackie Chen"
format: docx
---

[Forked repository](https://github.com/jackiechen24/chen-jackie_homework-03)

# Set Up

## Reading in packages/data
```{r}
# general use
library(tidyverse)
library(readxl)
library(here)
library(janitor)

# visualizing pairs
library(GGally)

# model selection
library(MuMIn)

# model predictions
library(ggeffects)

# model tables
library(gtsummary)
library(flextable)
library(modelsummary)

drought_exp <- read_xlsx(path = here("data", 
                                     "Valliere_etal_EcoApps_Data.xlsx"),
                         sheet = "First Harvest")

# quick look at data 
str(drought_exp)
class(drought_exp) # table dataframe object
```

## Cleaning
```{r}
# cleaning the data
drought_exp_clean <- drought_exp %>% 
  clean_names() %>% # nicer column names
  mutate(species_name = case_when( # adding column with species scientific names
    species == "ENCCAL" ~ "Encelia californica", # bush sunflower
    species == "ESCCAL" ~ "Eschscholzia californica", # California poppy
    species == "PENCEN" ~ "Penstemon centranthifolius", # Scarlet bugler
    species == "GRICAM" ~ "Grindelia camporum", # great valley gumweed
    species == "SALLEU" ~ "Salvia leucophylla", # Purple sage
    species == "STIPUL" ~ "Nasella pulchra", # Purple needlegrass
    species == "LOTSCO" ~ "Acmispon glaber" # deerweed
  )) %>% 
  relocate(species_name, .after = species) %>% # moving species_name column after species
  mutate(water_treatment = case_when( # adding column with full treatment names
    water == "WW" ~ "Well watered",
    water == "DS" ~ "Drought stressed"
  )) %>% 
  relocate(water_treatment, .after = water) # moving water_treatment column after water
```

# Problem 1. Multiple linear regression: model selection and construction (52 points)

## a. Make a table or list of all the models from class and the last one you constructed on your own. Write a caption for your table. (8 points)

1. **total mass is not predicted by any of the 3 predictors** (null model)

```{r}
model1 <- lm(total_g ~ 1, # formula
             data = drought_exp_clean) # data frame
```

2. **total mass predicted by SLA, water treatment, and species** (saturated model)

```{r}
# saturated model
model2 <- lm(total_g ~ sla + water_treatment + species_name,
             data = drought_exp_clean)
```

3. **total biomass as a function of SLA and water treatment**  

```{r}
model3 <- lm(total_g ~ sla + water_treatment,
             data = drought_exp_clean)
```

4. **total biomass as a function of SLA and species**  

```{r}
model4 <- lm(total_g ~ sla + species_name,
             data = drought_exp_clean)
```

5. **total biomass as a function of water treatment and species**

```{r}
model5 <- lm(total_g ~ water_treatment + species_name,
             data = drought_exp_clean)
```

Model selection
```{r}
model.sel(model1,
          model2, 
          model3, 
          model4,
          model5)
```

model5 (the one that includes the predictors of water treatment and species) has lowest AIC

Model diagnostics
```{r}
par(mfrow = c(2, 2))
plot(model5)
```

